WrapITK: Enhanced languages support\\ for the Insight Toolkit

### 2006-03-24 article/Article.tex
additions from Zach and Benoit
darcs-hash:20060324105919-2fc9d-b984f2f8aabd69fddc872c0925c10e98c7e36b50.gz

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann{$^1$}{\small{,}} Zachary Pincus{$^2$} {\small{and}} Benoit Regrain{$^3$}}
\authoraddress{{$^1$}Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France\\
{$^2$}Stanford University, CA 94305, USA\\
{$^3$}CREATIS, CNRS UMR 5515, 69621 Villeurbanne, France}

@@

@@

Currently, WrapITK have been tested on Mandriva Linux, and MacOS X, and only
supports Python properly; Java and Tcl libraries build
correctly but the Java and Tcl support classes for loading these libraries needs
some updating.

@@

@@

Here is the list of itk classes which can currently be substituted by python sequences:
\begin{itemize}
  \item \verb$CovariantVector$
  \item \verb$FixedArray$
  \item \verb$Index$
  \item \verb$Offset$
  \item \verb$Size$
  \item \verb$Vector$
\end{itemize}

@@

Some convenient functions are provided with the itk module. They all begin with a lower
case character to clearly show they are not part of ITK.
\begin{itemize}
  \item \verb$itk.image(object)$ try to return an image from the object given in
parameter. This function is used in most of the next functions to allow the user to
pass an image or a filter.

@@

@@

Several steps are necessary to obtain this nirvana, however. First, when
configuring the build in ccmake, you must set \verb$DOXYGEN_MAN_PATH$ to some directory
where man pages for the ITK classes will be created. Then, after the build, you
must run \verb$make_doxygen_config.py$ from within the \verb$Python$ directory in the build
directory, to collect information about the wrapped classes and create a doxygen
configuration file to make these man pages. Finally, run doxygen with that
configuration file. After these three simple steps, class docstrings will
contain the man page information. Note that this is limited to systems which
support the python \verb$commands$ module, and which have \verb$groff$ in the path. This
basically means anything but windows will work. (Cygwin should work too.)

@@

\verb$WRAP_CLASS$ issues an implicit call to \verb$WRAP_INCLUDE("ClassName.h")$, so the header
for the wrapped class itself does not need to be manually included. To disable
this behavior, set \verb$WRAPPER_AUTO_INCLUDE_HEADERS$ to \verb$OFF$.

The final optional parameter to \verb$WRAP_CLASS$ is \verb$POINTER$ or
\verb$POINTER_WITH_SUPERCLASS$. If no options are passed, then the class is wrapped
as-is. If \verb$POINTER$ is passed, then the class and the typedef'd \verb$class::Pointer$
type is wrapped. (\verb$Class::Pointer$ had better be a \verb$SmartPointer$ instantiation, or
things won't work. This is always the case for ITK-style code.) If
\verb$POINTER_WITH_SUPERCLASS$ is provided, then \verb$class::Pointer$, \verb$class::Superclass$ and
\verb$class::Superclass::Pointer$ are all wrapped. (Again, this only works for
ITK-style code where the class has a typedef'd \verb$Superclass$, and the superclass
has \verb$Self$ and \verb$Pointer$ typedefs.)

@@

  \item \verb$WRAPPER_AUTO_INCLUDE_HEADERS$. This variable is set to \verb$ON$ by default, but can
be set to \verb$OFF$ to disable the auto include feature. This feature should be used when several
classes to wrap come from the same header file.

@@

@@

  \item \verb$WRAP_IMAGE_FILTER(param_types param_count)$ is a more general
macro for wrapping image filters that need one or more image parameters of the
same type. The first parameter to this macro is a list of image pixel types for
which filter instantiations should be created. The second is a \verb$param_count$
parameter which controls how many image template parameters are created (see
above). The optional third parameter is a dimensionality condition (see above
also).

@@

    \subsubsection{Building}
To build an external project, first ensure that WrapITK has been properly built.
Then use ccmake to configure a build directory for the external project. If
WrapITK has not been installed, you will have to manually enter the path to the
WrapITK build directory.

@@

@@

I also thank Richard Beare for his early interest in using itk with python, and for
his useful feedbacks.

We thank Charl P. Botha for is help to debug WrapITK on windows plateform, and for the
patches he has contributed.

>>>>

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann{$^1$}{\small{,}} Zachary Pincus{$^2$} {\small{and}} Benoit Regrain{$^3$}}
\authoraddress{{$^1$}Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France\\
{$^2$}Program in Biomedical Informatics and Department of Biochemistry, Stanford University School of Medicine, Stanford, California\\
{$^3$}CREATIS, CNRS UMR 5515, 69621 Villeurbanne, France}

@@

  \item ITK was broken on MacOS X \cite{MacOsXWebSite} with python.

@@

Currently, WrapITK have been tested on Mandriva Linux \cite{MandrivaWebSite}, and MacOS X, and only
supports Python properly; Java and Tcl libraries build
correctly but the Java and Tcl support classes for loading these libraries is
entirely out of date and is not working.

@@

On linux boxes however, the user have to set the \verb$LD_LIBRARY_PATH$ to
point to libSwigRuntime.so. For example
\verb$export LD_LIBRARY_PATH=/usr/lib/InsightToolkit/WrapITK/Python-SWIG$.

@@

Here is the list of itk classes which can currently be substituted by python sequences:
\begin{itemize}
  \item \verb$CovariantVector$
  \item \verb$FixedArray$
  \item \verb$Index$
  \item \verb$Offset$
  \item \verb$Size$
  \item \verb$Vector$
  \item \verb$ContinuousIndex$
\end{itemize}

@@

Some convenient functions are provided with the itk module. They all begin with a lower
case character to clearly show they are not part of ITK.
\begin{itemize}
  \item \verb$itk.image(object)$ try to return an image from the object given in
parameter. If the object is an image, it is returned without changes. If the object
is a filter, the object returned by \verb$GetOutput()$ metyhod is returned.
This function is used in most of the next functions to allow the user to
pass an image or a filter, and is available here for the same usage in some custom
fuctions.

@@

This function is only a convenient function for a common task while prototyping.

@@

Several steps are necessary to obtain this nirvana, however. First, when
configuring the build in ccmake, you must set \verb$DOXYGEN_MAN_PATH$ to some directory
where man pages for the ITK classes will be created. Then, after the build, you
must run \verb$make_doxygen_config.py$ from within the \verb$Python$ directory in the build
directory, to collect information about the wrapped classes and create a doxygen
configuration file to make these man pages. Finally, run doxygen with that
configuration file. After these three simple steps, class docstrings will
contain the man page information. Note that this is limited to systems which
support the python \verb$commands$ module, and which have \verb$groff$ in the path. This
basically means anything but windows \cite{WindowsWebSite} will work. (Cygwin should work too.)

@@

\verb$WRAP_CLASS("itk::ImageFilter")$ issues an implicit call to \verb$WRAP_INCLUDE("itkImageFilter.h")$, so the header
for the wrapped class itself does not need to be manually included. To disable
this behavior, set \verb$WRAPPER_AUTO_INCLUDE_HEADERS$ to \verb$OFF$.

The final optional parameter to \verb$WRAP_CLASS$ is \verb$POINTER$ or
\verb$POINTER_WITH_SUPERCLASS$. If no options are passed, then the class is wrapped
as-is. If \verb$POINTER$ is passed, then the class and the typedef'd \verb$class::Pointer$
type is wrapped. (\verb$Class::Pointer$ had better be a \verb$SmartPointer$ instantiation, or
things won't work. This is always the case for ITK-style code.) If
\verb$POINTER_WITH_SUPERCLASS$ is provided, then \verb$class::Pointer$, \verb$class::Superclass$ and
\verb$class::Superclass::Pointer$ are all wrapped. (Again, this only works for
ITK-style code where the class has a typedef'd \verb$Superclass$, and the superclass
has \verb$Self$ and \verb$Pointer$ typedefs). \verb$POINTER_WITH_SUPERCLASS$ is especially
useful for wrapping classes whose superclasses depend on the template definitions of the given
filter. E.g. any of the functor image filters, which define  totally different superclass
template parameters depending on which functor is used.

@@

  \item \verb$WRAPPER_AUTO_INCLUDE_HEADERS$. This variable is set to \verb$ON$ by default, but can
be set to \verb$OFF$ to disable the auto include feature. This feature should be used when several
classes to wrap come from the same header file.
\verb$WRAPPER_AUTO_INCLUDE_HEADERS$ is re-set to \verb$ON$ for each new \verb$wrap_xxx.cmake$ file.

@@

These macros often take an optional second parameter which is a "dimensionality
condition" to restrict the dimensions that theis filter will be instantiated
for. The condition can either be a single number indicating the one dimension
allowed, a list of dimensions that are allowed (either as a single \verb$-$delimited
string or just a set of separate parameters), or something of the form \verb$n+$
(where \verb$n$ is a number) indicating that instantiations are allowed for dimension
n and above.

@@

  \item \verb$WRAP_IMAGE_FILTER(param_types param_count)$ is a more general
macro for wrapping image filters that need one or more image parameters of the
same type. The first parameter to this macro is a list of image pixel types for
which filter instantiations should be created. The second is a \verb$param_count$
parameter which controls how many image template parameters are created. The
optional third parameter is a dimensionality condition.

@@

    \subsubsection{Building}
To build an external project, first ensure that WrapITK has been properly built.
Then use \verb$ccmake$ to configure a build directory for the external project. If
WrapITK has not been installed, you will have to manually enter the path to the
WrapITK build directory.

By default, the build options are the same than the one used for building WrapITK,
but can be modified in the advanced options.

@@

ItkVtkGlue wrap the classes used to convert data from ITK to VTK \cite{VtkWebSite} and from VTK to ITK.
Those classes comes from the InsightApplications \cite{ITKWebSite}, and make the conversion
as simple in python as in C++. It has been tested with VTK 5.0.0.

By default, those classes are not loaded in the \verb$itk$ module to avoid loading
all the \verb$vtk$ module while importing \verb$itk$. The command \verb$import itkvtk$
will load the module, and add the classes both in \verb$itkvtk$ and\verb$itk$ module.
The classes are then usable as the other itk classes.

2> import itkvtk

3> reader = itk.ImageFileReader.IUC3.New()

4> converter = itk.ImageToVTKImageFilter.IUC3.New(reader)

5> converter.GetOutput()
5> <libvtkFilteringPython.vtkImageData vtkobject at 0xb7675b60>
\end{verbatim}

The \verb$itk.show3D$ class use the class \verb$ImageToVTKImageFilter$ to create the
volume rendering shown in Figure \ref{screenshot}.

@@

We would like to thank Charl P. Botha for is help to debug WrapITK on windows
plateform, and for the patches he has contributed, as well as Richard Beare for
his early interest in using itk with python, for his useful feedbacks, and for
his work on buffer conversion in python.

We thank Brad King for his assistance during the development process.
